---
title: "Access The Cancer Imaging Archive (TCIA) API with R"
author: "Marcel Ramos"
date: "2025-03-06"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 3
package: TCIAAPI
vignette: >
  %\VignetteIndexEntry{TCIAAPI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```


# TCIAAPI R package

## Installation

Currently, the package only resides on GitHub. 

```{r,eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("billila/TCIAAPI")
```

## Package load

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(TCIAAPI)
```

## Introduction

The `TCIAAPI` package provides an interface to the Cancer Imaging Archive (TCIA)
API. The TCIA API allows users to programmatically access the TCIA data. The
package provides functions to obtain an access token, download SVS images, and
retrieve metadata from the TCIA API.

## Access Token

The TCIA API requires an access token to access the data. The
`tcia_access_token` function retrieves the access token from the TCIA API. 
By default, it is configured to obtain a public token. Note that the token
expires after a certain period of time and must be refreshed.

```{r}
tcia_access_token() |> httr2::obfuscate()
```

Note that we use `httr2::obfuscate` to hide the token from the output.

## Information on SVS Images

The `tcia_svs_info` function retrieves metadata information on SVS images from
the TCIA API. The function requires a `camic_id` which is obtained from the
'TCIA Histopathology Custom Dataset Builder.json' file. The json file can be
obtained by navigating to the TCIA website
<https://www.cancerimagingarchive.net/> under 'Access The Data', 'Search
Histopathology Portal' and clicking on the 'TCIA Histopathology Custom Dataset
Builder' link.

```{r}
svsinfo <- tcia_svs_info("311781") 
svsinfo |> head(3L)
```

### Download URL

The `tcia_svs_info` function returns a list containing the metadata of the SVS
including the download URL. The download URL can be used to download the SVS
images. 

```{r}
svsinfo[["field_wsiimage"]][[1L]][["url"]]
```

Note that currently the package does not provide a function to download the ~150
MB json file programmatically.

## Download SVS Images

The `tcia_svs_download` function downloads SVS images from the TCIA API. Like
`tcia_svs_info`, the function requires a `camic_id` which can be obtained from
the 'TCIA Histopathology Custom Dataset Builder.json' file.

```{r,eval=FALSE}
tcia_svs_download("311781")
#' [1] "/tmp/RtmpjrBN7Z/MSB-02960-04-02.svs"
```

The function downloads the SVS images to the temporary directory by default. The
`destdir` argument can be used to specify a different directory.

# SessionInfo

```{r}
sessionInfo()
```

